\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-20 notes/2loopNLSM.tex
Standardized 'semi-abelian'

We can achieve this construction by splitting the gauge group of the full Yang-Mills action from $G$ to $G_L \times G_R$, where the full non-abelian gauge symmetry is restored only when we identify $G_L = G_R$. However, if we break the left gauge group down to $G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to $G_R \rightarrow U(N)$, then we are left with a semi-abelian theory that is gauge invariant in the preferred $U(1)$-directions. We call this manifestly color-dual theory, \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c .
\end{equation}
Here we can think of the abelian vector as a background field that sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix \cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the heart of YZ-theory \cite{Cheung:2016prv}, J-theory \cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills \cite{Monteiro2011pc}.
Specifically, semi-Abelian YM is simply a clever reinterpretation of the Lagrangian obtained by integrating an auxiliary field into the J-theory equations of motion.
The Feynman rule for the cubic vertex is
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
The propagator kinetic mixes between the two gauge fields and has the standard Feynman rule of $i/k^2$.
We can use the Ward identity of \cref{eq:GIrelA} to check that the amplitudes are indeed gauge invariant when the abelian polarizations are taken to be longitudinal, $\bar{\varepsilon} \rightarrow k$.

@@

\section{Conclusions and Outlook}\label{conclusions}
There is overwhelming evidence that YM and NLSM obey color-kinematics duality at loop level even though naive Lagrangians for these theories do not manifest the duality.
At tree level, the duality can be manifested by breaking manifest Bose symmetry at the Lagrangian level, either by explicitly picking different gluon states (as in SDYM) or by picking states that only restore Bose symmetry in the final amplitude (as in NLSM).
All of these theories can be encapsulated in the semi-Abelian YM theory presented in \cref{eq:semiYM} which obscures the Bose symmetry of the external states due to the presence of the two gauge fields $A_\mu$ and $\bar{A}_\mu$.
The kinetic mixing between the two gauge fields permits globally color-dual 1-loop integrands, but special care must be taken so that the integrands produced from semi-Abelian YM correspond to SDYM or NLSM.
Due to the kinetic mixing that stems from the broken Bose symmetry, it is impossible to construct a 2-loop integrand for semi-Abelian YM theory, that is, the theory is 1-loop exact.
Obtaining a globally color-dual integrand beyond 1-loop then requires sacrificing more than just Bose symmetry and in this paper it is argued that off-shell locality must be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in order to obtain color-dual 2-loop integrands.}

@@

Below we review the four manifestly color-dual theories that appear in this paper, in particular SDYM, two formulations of NLSM, and ZM.
All of these theories are closely tied together by the semi-Abelian YM theory in \cref{eq:semiYM} or in other words the cubic sector of YM.
The manifestly color-dual equations of motion for SDYM and NLSM are derived by picking a gauge and then placing a constraint not on the YM equation of motion or Bianchi identity but rather directly on the field strength itself.
Other theories can be obtained from SDYM and NLSM by judicious modification of the propagator.
In all of these theories, the kinematic algebra is that of (volume preserving) diffeomorphisms or some close variant thereof.
In almost all of these theories, Bose symmetry is broken or obscured at the Lagrangian level.
Because of the broken Bose symmetry (and the inherited kinetic mixing of semi-Abelian YM), these theories cannot produce color-dual numerators at 2-loop or higher.

@@

\paragraph{J-theory} There is an alternative color-dual formulation of NLSM that the theory manifestly corresponds to NLSM at least at tree level \cite{Cheung:2021zvb}.
For related formulations of NLSM that do not treat color-kinematics duality see \cite{Freedman:1980us, Slavnov:1971mz}.
The construction begins with a first order formulation of NLSM in terms of chiral currents $J^\mu$.
First order formulation of NLSM in terms of chiral current $J^\mu$.
Two equations.
Two constraints, the field strength should vanish
\begin{equation}
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0
\end{equation}
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge
\begin{equation}
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0
\end{equation}
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\begin{equation}
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\end{equation}
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-Abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
In other words, \emph{J-theory and semi-Abelian YM are identical}.
By calculating an off-shell 4-point correlation function it is possible to check that this Lagrangian automatically satisfies color-kinematics duality.
As might be expected, the kinematic algebra is the diffeomorphism algebra.
For details about external states and how to extract NLSM (a scalar theory) from semi-YM (a vector theory) we refer the reader to \cite{Cheung:2021zvb}.
Note that the different external particles break manifest Bose symmetry.

>>>>

We can achieve this construction by splitting the gauge group of the full Yang-Mills action from $G$ to $G_L \times G_R$, where the full non-abelian gauge symmetry is restored only when we identify $G_L = G_R$. However, if we break the left gauge group down to $G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to $G_R \rightarrow U(N)$, then we are left with a semi-abelian theory that is gauge invariant in the preferred $U(1)$-directions. We call this manifestly color-dual theory, \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c .
\end{equation}
Here we can think of the abelian vector as a background field that sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix \cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the heart of YZ-theory \cite{Cheung:2016prv}, J-theory \cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills \cite{Monteiro2011pc}.
Specifically, semi-abelian YM is simply a clever reinterpretation of the Lagrangian obtained by integrating an auxiliary field into the J-theory equations of motion.
The Feynman rule for the cubic vertex is
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
The propagator kinetic mixes between the two gauge fields and has the standard Feynman rule of $i/k^2$.
We can use the Ward identity of \cref{eq:GIrelA} to check that the amplitudes are indeed gauge invariant when the abelian polarizations are taken to be longitudinal, $\bar{\varepsilon} \rightarrow k$.

@@

\section{Conclusions and Outlook}\label{conclusions}
There is overwhelming evidence that YM and NLSM obey color-kinematics duality at loop level even though naive Lagrangians for these theories do not manifest the duality.
At tree level, the duality can be manifested by breaking manifest Bose symmetry at the Lagrangian level, either by explicitly picking different gluon states (as in SDYM) or by picking states that only restore Bose symmetry in the final amplitude (as in NLSM).
All of these theories can be encapsulated in the semi-abelian YM theory presented in \cref{eq:semiYM} which obscures the Bose symmetry of the external states due to the presence of the two gauge fields $A_\mu$ and $\bar{A}_\mu$.
The kinetic mixing between the two gauge fields permits globally color-dual 1-loop integrands, but special care must be taken so that the integrands produced from semi-abelian YM correspond to SDYM or NLSM.
Due to the kinetic mixing that stems from the broken Bose symmetry, it is impossible to construct a 2-loop integrand for semi-abelian YM theory, that is, the theory is 1-loop exact.
Obtaining a globally color-dual integrand beyond 1-loop then requires sacrificing more than just Bose symmetry and in this paper it is argued that off-shell locality must be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in order to obtain color-dual 2-loop integrands.}

@@

Below we review the four manifestly color-dual theories that appear in this paper, in particular SDYM, two formulations of NLSM, and ZM.
All of these theories are closely tied together by the semi-abelian YM theory in \cref{eq:semiYM} or in other words the cubic sector of YM.
The manifestly color-dual equations of motion for SDYM and NLSM are derived by picking a gauge and then placing a constraint not on the YM equation of motion or Bianchi identity but rather directly on the field strength itself.
Other theories can be obtained from SDYM and NLSM by judicious modification of the propagator.
In all of these theories, the kinematic algebra is that of (volume preserving) diffeomorphisms or some close variant thereof.
In almost all of these theories, Bose symmetry is broken or obscured at the Lagrangian level.
Because of the broken Bose symmetry (and the inherited kinetic mixing of semi-abelian YM), these theories cannot produce color-dual numerators at 2-loop or higher.

@@

\paragraph{J-theory} There is an alternative color-dual formulation of NLSM that the theory manifestly corresponds to NLSM at least at tree level \cite{Cheung:2021zvb}.
For related formulations of NLSM that do not treat color-kinematics duality see \cite{Freedman:1980us, Slavnov:1971mz}.
The construction begins with a first order formulation of NLSM in terms of chiral currents $J^\mu$.
First order formulation of NLSM in terms of chiral current $J^\mu$.
Two equations.
Two constraints, the field strength should vanish
\begin{equation}
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0
\end{equation}
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge
\begin{equation}
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0
\end{equation}
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\begin{equation}
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\end{equation}
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
In other words, \emph{J-theory and semi-abelian YM are identical}.
By calculating an off-shell 4-point correlation function it is possible to check that this Lagrangian automatically satisfies color-kinematics duality.
As might be expected, the kinematic algebra is the diffeomorphism algebra.
For details about external states and how to extract NLSM (a scalar theory) from semi-YM (a vector theory) we refer the reader to \cite{Cheung:2021zvb}.
Note that the different external particles break manifest Bose symmetry.
